Empirical Analysis of the Role of Energy in Economic Growth

### 2013-01-30 EconGrowth.Rnw
Now gathering predictions for the US from the CD model.

<<Country Factors Graph, echo=FALSE>>=
createCountryFactorsGraph <- function(countryName){
  dataTable <- loadData(countryName)
  graphType <- "l"
  lineTypes <- c(1, 5, 2, 4, 6) #line types. See http://en.wikibooks.org/wiki/R_Programming/Graphics
  lineWidths <- c(1, 1, 1, 1, 1) #line widths
  colors <- c("black", "black", "black", "black", "black") #line and point colors
  lineSpec <- list(lty=lineTypes, lwd=lineWidths, col=colors)
  graph <- xyplot(iCapStk+iLabor+iQ+iX+iU ~ Year, data=dataTable,
                  type=graphType,
                  par.settings = list(superpose.line = lineSpec),
                  key=list(text=list(c("k", "l", "q", "x", "u")),
                           type=graphType,
                           lines=lineSpec,
                           columns=1, x=0.0, y=0.98),
                  scales=list(cex=1.0, #controls text size on scales
                              tck=-0.5), #controls tick mark length
                  ylab="Indexed (1980=1)")
  return(graph)
}
@
<<Factors Lattice Graph, fig.width=6.5, fig.height=7.5, fig.cap="GDP ($y$), capital stock ($k$), labor ($l$), thermal energy ($q$), exergy ($x$), and useful work ($u$) for all economies. (China's indexed GDP and indexed capital stock rise to $y = 7.3$ and $k = 9.2$ in 2011.)", echo=FALSE>>=
createFactorsLatticeGraph <- function(countryName){
  dataTable <- loadData(countryName)
  graphType <- "b"
  lineTypes <- c(0, 1, 5, 2, 4, 1) #line types. See http://en.wikibooks.org/wiki/R_Programming/Graphics
  lineWidths <- c(0, 2, 2, 1, 1, 1) #line widths. 0 means no line.
  colors <- c("black", "black", "black", "red", "blue", "black") #line and point colors
  symbols <- c(1, NA_integer_, NA_integer_, NA_integer_, NA_integer_, NA_integer_)  #NA_integer_ gives no symbol.
  graph <- xyplot(iGDP+iCapStk+iLabor+iQ+iX+iU ~ Year | Country, data=dataTable,
                  type=graphType,
                  index.cond=list(c(7,6,3,1,8,5,4,2,9)), #orders the panels. Default is alphabetical.
                  layout=c(3,3), #indicates a 3x3 arrangement of panels.
                  as.table=TRUE, #indexing of panels starts in upper left and goes across rows.
                  lty = lineTypes, lwd = lineWidths, col = colors, #Controls line parameters
                  pch = symbols, col.symbol = colors, #Controls symbol parameters
                  key=list(text=list(c("y", "k", "l", "q", "x", "u")),
                           type=graphType,
                           cex=0.85, #controls size of text in the key
                           lines=list(lty=lineTypes, lwd=lineWidths), #controls line types
                           pch=symbols, col=colors, #controls symbol (plot characters) types
                           columns=1, x=0.01, y=0.95), #We want only one column and sets location of the key
                  scales=list(cex=0.75, #controls text size on scales.
                              tck=-0.5, #controls tick mark length. < 0 for inside the graph.
                              y=list(at=c(1,2,3,4,5,6,7,8,9,10))), #tick mark locations
                  ylim=c(0.5,4.5), #y axis limits
                  xlab="", ylab="Indexed (1980=1 or 1991=1)") #axis labels
  return(graph)
}
createFactorsLatticeGraph("All") #Reads data from the "AllData.txt" file.
@

@@

<<cobb-douglas, echo=FALSE>>=

#################################################
# Calculates parameter estimates and confidence intervals
# for the Cobb-Douglas production function given a country.
#
# countryName is a string containing the 2-letter abbreviation for the country, e.g. "US" or "CN"
#
# returns a vector of data for the Cobb-Douglas model. 
# First item is the +95% CI on all parameters
# Second item contains the parameter estimates
# Third item is the -95% CI on all parameters
# Each row has names: lambda, alpha, beta, gamma, corresponding to the parameters in the model.
##
cobbDouglasData <- function(countryName){

@@

  # Runs a non-linear least squares fit to the data. We've replaced beta with 1-alpha for simplicity.
  modelCD <- nls(iGDP ~ exp(lambda*iYear) * iCapStk^alpha * iLabor^(1 - alpha), 
#                 algorithm = "port",
                 start = list(lambda=lambdaGuess, alpha=alphaGuess),
#                 lower = list(lambda=-Inf, alpha=0),
#                 upper = list(lambda=Inf, alpha=1),
                 data=dataTable)

@@

cobbDouglasParamsTable <- function(){
  dataCD <- rbind(cobbDouglasCountryRow("US"),
                cobbDouglasCountryRow("UK"),
                cobbDouglasCountryRow("JP"),
                cobbDouglasCountryRow("CN"),
                cobbDouglasCountryRow("ZA"),
                cobbDouglasCountryRow("SA"),
                cobbDouglasCountryRow("IR"),
                cobbDouglasCountryRow("TZ"),
                cobbDouglasCountryRow("ZM"))
  colnames(dataCD) <- c(" ", "$\\lambda$", " ", 
                        " ", "$\\alpha$",  " ",
                        " ", "$\\beta$",   " ")
  rownames(dataCD) <- countries
  tableCD <- xtable(dataCD, 
                    caption="Cobb-Douglas parameters for 1980-2011 (US, UK, JP) or 1991-2011 (others). (Parameter estimates beneath symbol. 95\\% confidence bounds to left and right.)", 
                    digit = c(1, 4,4,4, 2,2,2, 2,2,2),
                    align = "r|ccc|ccc|ccc")
  return(tableCD)
}
@

@@

>>>>

<<Country Factors Graph, echo=FALSE>>=
createCountryFactorsGraph <- function(countryName){
  dataTable <- loadData(countryName)
  graphType <- "l"
  lineTypes <- c(1, 5, 2, 4, 6) #line types. See http://en.wikibooks.org/wiki/R_Programming/Graphics
  lineWidths <- c(1, 1, 1, 1, 1) #line widths
  colors <- c("black", "black", "black", "black", "black") #line and point colors
  lineSpec <- list(lty=lineTypes, lwd=lineWidths, col=colors)
  graph <- xyplot(iCapStk+iLabor+iQ+iX+iU ~ Year, data=dataTable,
                  type=graphType,
                  par.settings = list(superpose.line = lineSpec),
                  key=list(text=list(c("k", "l", "q", "x", "u")),
                           type=graphType,
                           lines=lineSpec,
                           columns=1, x=0.0, y=0.98),
                  scales=list(cex=1.0, #controls text size on scales
                              tck=-0.5), #controls tick mark length
                  ylab="Indexed (1980=1)")
  return(graph)
}
@
<<Factors_Lattice_Graph, fig.width=6.5, fig.height=7.5, fig.cap="GDP ($y$), capital stock ($k$), labor ($l$), thermal energy ($q$), exergy ($x$), and useful work ($u$) for all economies. (China's indexed GDP and indexed capital stock rise to $y = 7.3$ and $k = 9.2$ in 2011.)", echo=FALSE>>=
createFactorsLatticeGraph <- function(countryName){
  dataTable <- loadData(countryName)
  graphType <- "b"
  lineTypes <- c(0, 1, 5, 2, 4, 1) #line types. See http://en.wikibooks.org/wiki/R_Programming/Graphics
  lineWidths <- c(0, 2, 2, 1, 1, 1) #line widths. 0 means no line.
  colors <- c("black", "black", "black", "red", "blue", "black") #line and point colors
  symbols <- c(1, NA_integer_, NA_integer_, NA_integer_, NA_integer_, NA_integer_)  #NA_integer_ gives no symbol.
  graph <- xyplot(iGDP+iCapStk+iLabor+iQ+iX+iU ~ Year | Country, data=dataTable,
                  type=graphType,
                  index.cond=list(c(7,6,3,1,8,5,4,2,9)), #orders the panels. Default is alphabetical.
                  layout=c(3,3), #indicates a 3x3 arrangement of panels.
                  as.table=TRUE, #indexing of panels starts in upper left and goes across rows.
                  lty = lineTypes, lwd = lineWidths, col = colors, #Controls line parameters
                  pch = symbols, col.symbol = colors, #Controls symbol parameters
                  key=list(text=list(c("y", "k", "l", "q", "x", "u")),
                           type=graphType,
                           cex=0.85, #controls size of text in the key
                           lines=list(lty=lineTypes, lwd=lineWidths), #controls line types
                           pch=symbols, col=colors, #controls symbol (plot characters) types
                           columns=1, x=0.01, y=0.95), #We want only one column and sets location of the key
                  scales=list(cex=0.75, #controls text size on scales.
                              tck=-0.5, #controls tick mark length. < 0 for inside the graph.
                              y=list(at=c(1,2,3,4,5,6,7,8,9,10))), #tick mark locations
                  ylim=c(0.5,4.5), #y axis limits
                  xlab="", ylab="Indexed (1980=1 or 1991=1)") #axis labels
  return(graph)
}
createFactorsLatticeGraph("All") #Reads data from the "AllData.txt" file.
@

@@

<<cobb-douglas, echo=FALSE>>=
####################
# Returns an nls Cobb-Douglas model for the country specified
##
cobbDouglasModel <- function(countryName){
  # Load the data that we need.
  dataTable <- loadData(countryName)

@@

  # Runs a non-linear least squares fit to the data. We've replaced beta with 1-alpha for simplicity.
  modelCD <- nls(iGDP ~ exp(lambda*iYear) * iCapStk^alpha * iLabor^(1 - alpha), 
#                 algorithm = "port",
                 start = list(lambda=lambdaGuess, alpha=alphaGuess),
#                 lower = list(lambda=-Inf, alpha=0),
#                 upper = list(lambda=Inf, alpha=1),
                 data=dataTable)
  return(modelCD)
}

#################################################
# Calculates parameter estimates and confidence intervals
# for the Cobb-Douglas production function given a country.
#
# countryName is a string containing the 2-letter abbreviation for the country, e.g. "US" or "CN"
#
# returns a vector of data for the Cobb-Douglas model. 
# First item is the +95% CI on all parameters
# Second item contains the parameter estimates
# Third item is the -95% CI on all parameters
# Each row has names: lambda, alpha, beta, gamma, corresponding to the parameters in the model.
##
cobbDouglasData <- function(countryName){
  modelCD <- cobbDouglasModel(countryName)

@@

############################
# Aggregates the Cobb-Douglas results into a big data table
##
cobbDouglasParamsTable <- function(){
  dataCD <- rbind(cobbDouglasCountryRow("US"),
                cobbDouglasCountryRow("UK"),
                cobbDouglasCountryRow("JP"),
                cobbDouglasCountryRow("CN"),
                cobbDouglasCountryRow("ZA"),
                cobbDouglasCountryRow("SA"),
                cobbDouglasCountryRow("IR"),
                cobbDouglasCountryRow("TZ"),
                cobbDouglasCountryRow("ZM"))
  colnames(dataCD) <- c(" ", "$\\lambda$", " ", 
                        " ", "$\\alpha$",  " ",
                        " ", "$\\beta$",   " ")
  rownames(dataCD) <- countries
  tableCD <- xtable(dataCD, 
                    caption="Cobb-Douglas parameters for 1980-2011 (US, UK, JP) or 1991-2011 (others). (Parameter estimates beneath symbol. 95\\% confidence bounds to left and right.)", 
                    digit = c(1, 4,4,4, 2,2,2, 2,2,2),
                    align = "r|ccc|ccc|ccc")
  return(tableCD)
}
@

@@

<<CD Lattice Graph>>=
usModel <- cobbDouglasModel("US")
coefs <- coef(usModel)
print(coef(usModel))
usPred <- predict(usModel)
class(usPred)
print(usPred)
data.frame(usPred)

@
